On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-08 manuscripts/IEEE/final_submission.tex
minor

We claim that a similar phenomenon applies to datasets of vectors with a stable sparse representation. Briefly, following the same procedure as in \cite[Sec.~IV]{Hillar15}, for $k < m$ and $n \geq \min(2k, m)$, we may consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by an indeterminate $n \times m$ matrix $\mathbf{A}$ and $m$-dimensional $k$-sparse vectors $\mathbf{x}_1, \ldots, \mathbf{x}_N$ indeterminate within their supports, which form a regular hypergraph $\mathcal{H} \subseteq {[m] \choose k}$ satisfying the SIP. Restricting \mbox{$(k-1){m \choose k} + 1$} indeterminate $\mathbf{x}_i$ to each support in $\mathcal{H}$, and letting $\textbf{M}$ be the $n \times N$ matrix with columns $\mathbf{A}\mathbf{x}_i$, it can be checked that when $f(\mathbf{M}) \neq 0$ for a substitution of real numbers for the indeterminates, all of the assumptions on $\mathbf{A}$ and $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied. We therefore have the following result.  % ; in particular, $\mathbf{A}$ satisfies \eqref{SparkCondition}

@@

We close this section with some comments on the optimality of our results.  The linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse coding model? Our results demonstrate that sparse codes $\mathbf{x}_i$ drawn from only a polynomial number of $k$-dimensional subspaces permit stable identification of the generating dictionary $\mathbf{A}$. This lends some legitimacy to the use of the sparse coding model in practice, where data in general are unlikely (if ever) to exhibit the exponentially many possible $k$-wise combinations of dictionary elements required by (to our knowledge) all previously published results. Consequently, if $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices to identify the dictionary.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, Tillmann15}. Even so, in the next section we discuss how probabilistic guarantees can in fact be made for any number of available samples.

>>>>

We claim that a similar phenomenon applies to datasets of vectors with a stable sparse representation. Briefly, following the same procedure as in \cite[Sec.~IV]{Hillar15}, for $k < m$ and $n \geq \min(2k, m)$, we may consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by an indeterminate $n \times m$ matrix $\mathbf{A}$ and $m$-dimensional $k$-sparse vectors $\mathbf{x}_1, \ldots, \mathbf{x}_N$ indeterminate within their supports, which form a regular hypergraph $\mathcal{H} \subseteq {[m] \choose k}$ satisfying the SIP. Restricting \mbox{$(k-1){m \choose k} + 1$} indeterminate $\mathbf{x}_i$ to each support in $\mathcal{H}$, and letting $\textbf{M}$ be the $n \times N$ matrix with columns $\mathbf{A}\mathbf{x}_i$, it can be checked that when $f(\mathbf{M}) \neq 0$ for a substitution of real numbers for the indeterminates, all of the assumptions on $\mathbf{A}$ and $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem} are satisfied. We therefore have the following.  % ; in particular, $\mathbf{A}$ satisfies \eqref{SparkCondition}

@@

We close this section with some comments on the optimality of our results.  The linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse coding model? Our results demonstrate that sparse codes $\mathbf{x}_i$ drawn from only a polynomial number of $k$-dimensional subspaces permit stable identification of the generating dictionary $\mathbf{A}$. This lends some legitimacy to the use of the model in practice, where data in general are unlikely (if ever) to exhibit the exponentially many possible $k$-wise combinations of dictionary elements required by (to our knowledge) all previously published results. Consequently, if $k$ is held fixed or if the size of the support set of reconstructing codes is polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices to identify the dictionary.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, Tillmann15}. Even so, in the next section we discuss how probabilistic guarantees can in fact be made for any number of available samples.
